A thesis draft

### 2018-12-03 res/sections/15-ArchitecturalImplementation.tex
Fix Arch impl

In our first attempt we started building our infrastructure on top of Ubuntu
16.04 (for Openstack) and Ubuntu 18.04 (for Openbaton). Our main goal was to set
up Openstack as our VIM while setting Openbaton as our MANO. Then, Openbaton
would have registered Openstack as PoP, where TOSCA definitions would have been
launched on Openstack that, instead of deploying hypervisor based virtual
machines, it would have deployed container instances. The VNFs lifecycles would
have been controlled through an Element Management System (EMS), transmitting
the various VNF states via the message broker RabbitMQ. It is possible to see the
interactions between elements in a minimum schema available in
Figure~\ref{chap:archimpl:sec:fistattempt:img:schema1}.

@@

Our first step was to try and use Openstack to create a container orchestration.
The reason behind this approach was simple: Openbaton, on the paper, comes with
an out-of-the-box support for Openstack. In fact, it natively supports Openstack
PoPs, giving to us a great possibility to reuse the two frameworks and to ease
the development phase of our software.
\subsubsection{Devstack: Openstack for developers}
Openstack is composed by modules, giving the possibility to the user who wants
to install it to choose only the components it really needs. So the first thing
we decided was to perform a selection of these modules between the 33 available
in the installation page. The installation of this components span between
multiple machines (nodes), therefore creating a whole cluster of resources.
Since we considered too time-consuming to install Openstack in a distributed
manner, we considered to test first the developer version, called
\emph{Devstack}. Devstack offers the possibility to install a subset of modules
in the local machine, and to leverage a Openstack environment able to perform
the same operations of the production-grade installation. Thus, we started
configuring the Devstack installation, installing first of all the Nova module,
that is required to perform computing operations such as the launch of virtual
machines, and lately we installed Tacker too.

@@

First of all, Devstack needs a machine with at least 16GB of ram and 50GB of HD 
space.
We decided to install these components:
\begin{itemize}
 \item Keystone (for identity management)
 \item Object storage
 \item Compute
 \item Tacker (we tested this module to see is Openstack was a suitable VIM too)
\end{itemize}

@@

A short description about every installation/deploy method will be briefly
illustrated.

@@

@@

After the initial Openstack and Openbaton configuration we started making the
components talking to each other. First, we tried a simple \verb!iperf! data
transmission with two VNFs deployed by Openbaton as a common hypervisor-based
Openstack instances, as explained in~\cite{openbatonIperf}. With this test
completed, we started tinkering Openstack for using Docker containers instead of
traditional hypervisor instances. When we started developing this solution, we
adopted the Zun module, that aims to smoothly integrate Docker in Openstack.
Unfortunately, we were not able to make it properly work, in particular, it was
not possible to open desired ports on the Docker containers. This lead to the
impossibility to perform packet forwarding between VNFs. On top of that, we
found difficulties regarding the integration of Docker images in TOSCA
definitions and Openstack. These two issues are related: the TOSCA definition
does not have any keyword to specificate in any way that the image used for the
VNF definition is a Docker image, and Openstack, with the Zun plugin, at the
time of the implementation did not performed the image pull from any Docker
registry. The impossibility to easily set the containers ports and the lack of
some support between the TOSCA definition and the Openstack Zun plugin pushed us
to drop Openstack as a possible solution and to start looking for another valid
VIM component.

@@

After the failure to find a viable setting for Openstack, we decided to look
around for other solutions. We started studying proper Docker container
orchestrators that could offer all the features we needed: container lifecycle
management, scaling, healing and monitoring. We decided to pick two of the most
mature and famous ones: Docker Swarm and Kubernetes. The decision to study only
these two relies on their big community supporting them and the possibility to
easily found projects that used the aforementioned frameworks.

@@

First we started playing with Docker Swarm. Docker Swarm is a product directly
made by the Docker Inc., so it seamlessly integrates with other Docker tools
such as Docker Compose. The installation and set-up is pretty straightforward,
and the configuration is almost automatically handled by Docker itself. The
minimum number of nodes in order to create a proper Docker Swarm cluster is of
three, where one must be a manager node (designed to orchestrate the other
nodes in the cluster), while the others can be simple workers. Multiple managers
can coexist in the same cluster, as long as only one effectively operates. The
others act like backup manager in case the one orchestrating goes down for some
reason. The networking is given out-of-the box, and the same holds for
distributed persistence. While this for most of the uses cases can be considered
a pro, in our specific situation the impossibility to tweak the network
configuration was a cons, since it was important for us to have the full control
over the network, especially regarding the SFC implementation, explained better
in Chapter~\ref{chap:vnf_ns_impl}.

\subsection{Kubernetes}
Kubernetes, as already described in the technology section, is an orchestrator
created by Google able to create a software abstraction from the underlying
virtualization technology. This enables Kubernetes to create an integration with
Docker. Kubernetes has an architecture based on plug-ins, giving the possibility
to the users to change part of Kubernetes itself, such as the networking routing
strategy or the persistence layer. With this flexibility Kubernetes offered to
us what Docker Swarm was not able to give. Additionally, we had the possibility
to set custom Domain Name System (DNS) lookups and distribute other internal
components in order to avoid requests bottleneck.

@@

Kubernetes does not have a minimum number of nodes in order to properly work.
Nonetheless, we decided to use a four node cluster configured in the following
way: three slaves nodes (also known as \emph{minions}) and one master node. The
Kubernetes version installed was the 1.11, enabling us to exploit new features
like a reworked DNS lookup system and new storage features\footnote{A full
  kubernetes log release can be found at:
  \url{https://kubernetes.io/blog/2018/06/27/kubernetes-1.11-release-announcement/}}.
We initially deployed from the internal Openstack of the University four nodes
with 4GB of RAM, 10 GB of SSD storage and 2 vCPUs on Ubuntu 16.04 machines (that
we later swapped with CentOS 7). We performed the installation using the
official Kubernetes repositories. After the initial installation, we obtained a
configuration similar to the one illustrated in
Figure~\ref{chap:archimpl:sec:secondattempt:img:k8scp}. In this figure is
possible to denote different components on the master node and on the minions,
which will be described.

@@

After the Kubernetes configuration and installation, we started working for a
viable integration between Kubernetes and Openbaton. Our first idea is the one
illustrated in Figure~\ref{chap:archimpl:sec:secondattempt:img:schema1}:
Openbaton, that usually handles the TOSCA definitions storing them in its
database, had to somehow convert these definition in a suitable YAML for
Kubernetes, that had in turn to elaborate it and schedule to one minion in the
cluster. The monitoring and management would remain untouched, as we planned to
create an image with the integrated EMS tools that provided a seamless
integration with Docker container. With this setting, the only thing to solve
was how to convert TOSCA definitions in YAML one. The first thing we started
thinking about was how to perform the conversion. This topic was quite important
because the TOSCA definition does not really consider the possibility to have an
alternative virtualization system apart from the hypervisor one. This causes the
TOSCA definition to be, in our opinion, too much focused on the resources usage
and granularity aspects such the deployment flavour and the kind of virtual link
(concept that does not really exists in container based systems) to instantiate.
Nonetheless, we performed a mapping between most of the TOSCA keywords with some
of the YAML definitions, in a way that a TOSCA configuration file was able to
generate a valid YAML deployment. After that, we started thinking instead who,
in the current architecture, had the task to perform this operation. Initially
we thought that Kubernetes itself could have been a perfect candidate. From a
received TOSCA, a Kubernetes component should have validated and translated the
TOSCA definition to generate a YAML suitable to be finally submitted to the
Kubernetes API Server. Another feasible approach could have been to create a
Openbaton plugin in order to convert the TOSCA definition before submitting it
to Kubernetes for the deployment. This solution had great advantages in terms of
responsibility separation: in this way only the MANO had to know, for a VIM
based on Kubernetes, that a conversion was needed. From a Kubernetes
point-of-view, nothing would have changed, since all the configuration
conversion would have been performed by Openbaton.

@@

The impossibility to develop a custom VIM driver and VNFM manager led us to
search for an alternative solution. Then, we started to work directly on the
VIM. The base approach was to create an API Server which in future an Openbaton
plugin could contact and simply deploy TOSCA definitions, interacting with a set
of RESTful APIs. We called this simple backend \emph{Harbor}.

@@

Terminated the Harbor development, we ended up in a situation similar the one
depicted in Figure~\ref{chap:archimple:sec:secondattempt:img:attempt2v1harbor}.
It is possible to see how the Openbaton framework is not necessary anymore:
Harbor, in fact, ended up substituting it. We then started to think about the
challenges of integrating the SFC structure inside a Kubernetes cluster. We
identified the following challenges:
\begin{itemize}
\item whole system scalability: we wanted that every component had the
  possibility to scale during intense workloads, without the necessity of human
  intervention.
\item ingress and egress points: there should be at least two points in the
  cluster were the packet enters and exits;
\item dynamic package routing: as formulated by one of our requirements, packet
  routing had to be dynamic. We immediately identified a possibility to create
  this functionality with the already-built features of Kubernetes;
\item a framework to automatically forward the data elaborated by the VNF: we
  identified the possibility to build a framework that automatically handled the
  packet receiving from another VNF, it properly called the function that had to
  elaborate the packet payload and then it forwarded the result to the next VNF;
\item connection status: having ingress and egress points implicates the
  necessity to save some persistent data in a distributed manner.
\end{itemize}

\subsubsection{Scalability considerations}
When adopting Kubernetes, we realized the possibility to make every component
easily scalable. Indeed, all the following components can be described in
Kubernetes as a \verb!Service!. For a \verb!Service! a consequent
\verb!Deployment! can be associated, giving the possibility for it to scale
under heavy workloads. The introduction of scalability meant that it was not
possible to refer to the components based on their IP, but we had to refer to
their service name. As it will be explained later, this introduced challenges
especially for the ingress and egress points of our network, that have to keep a
connection with the sender and the receiver. On top of that, the scalability and
the required DNS lookups meant a possible bottleneck. With a great number of
components and VNFs constantly querying the centralized Kubernetes DNS lookup
system a performance degradation could have easily happened. We decided to
distribute the DNS service too, integrating the CoreDNS Kubernetes default
service with \emph{Dnsmasq}, a DNS server able to perform query caching for
faster name resolution. We created a Dnsmasq \verb!DaemonSets!, which in turn
deployed one Pod with Dnsmasq running inside for every minion in the cluster. At
the time of the VNF launch, we set that the DNS resolution should pass first for
the local Dnsmasq DNS, and then for the Kubernetes CoreDNS service. In this way,
we were able to distribute the DNS query workload from the master, using a local
caching system to speed up the name resolution between services.

\subsubsection{Ingress and Egress: Ironhide}
As just said, the whole framework needs to have and entry and an exit point.
These, defined as \emph{Ingress} and \emph{Egress} points, are performed by
\emph{Ironhide}. Ironhide hides the whole packet processing to the sender $S$
and the receiver $R$. When a client $S$ performs a request to $R$, the packet
actually goes through Ironhide, who has an internal classifier that it is able
to recognize the type of connection and to choose the right NS the packet has to
go through. This process is fundamental since it keeps the connection between
$S$ and $R$ up, for both TCP and UDP based communications. Since the connection
needs to be end-to-end, the state of the connections are saved in the
\emph{Roulette Controller} component.

@@

\subsubsection{Keeping the connection status}
The scalability challenge introduces new issues in the system. The most
important is keeping up stateful connections, since the ingress and egress
points could be more than one container. With this consideration in mind, we
started thinking at how and where keeping information about the ingress
container handling the connection with the sender $S$ and about the egress
container talking with the receiver $R$. Since the Roulette component has a
complete end-to-end topology view of the system, we decided to store these
additional information (we called it \emph{endpoints}) in it.

@@

\subsection{Discussion}
At the end, we were able to create a scalable architecture that followed the
ETSI MANO standard, with Kubernetes as VIM and Harbor has MANO. We faced
additional challenges regarding the system scalability, that we solved using the
Roulette controller. We eased the creation of VNFs with Astaire, and we created
proper ingress and egress points with Ironhide, that supported scalability
issues and performed a protocol-based packet classification. Is possible to see
a final architectural schema in
Figure~\ref{chap:archimpl:sec:secondattempt:img:attempt2v1total}.

>>>>

In our first attempt we started building our infrastructure on top of Ubuntu
16.04 (for Openstack) and Ubuntu 18.04 (for Openbaton). Our main goal was to set
up Openstack as our VIM while setting Openbaton as our MANO. Then, Openstack
would register with Openbaton and TOSCA definitions would be deployed on
Openstack following the Infrastructure as Code (IaC) paradigm. In this context
we are interested in deploying container-based instances instead of hypervisor
based virtual machines. The VNFs lifecycles would have been controlled through
an Element Management System (EMS), transmitting the various VNF state changes
via the message broker RabbitMQ. It is possible to see the interactions between
the elements in a minimal schema available in
Figure~\ref{chap:archimpl:sec:fistattempt:img:schema1}.

@@

Our first step was to try and use Openstack to create a container orchestration
environment. The reason behind this approach was simple: Openbaton, on the
paper, comes with an out-of-the-box support for Openstack. In fact, it natively
supports Openstack registered as a PoP, giving to us a great possibility to
reuse the two frameworks and to ease the development phase of our software.
\subsubsection{Devstack: Openstack for developers}
Openstack is composed by modules, giving the possibility to the user who wants
to install it to choose only the components it really needs. So the first thing
we decided was to perform a selection of these modules between the 33 available
in the installation page. The installation of this components span between
multiple machines (nodes), therefore creating a whole cluster of resources.

Since a distributed Openstack installation was too time consuming, we considered
to test first the developer version, called \emph{Devstack}. Devstack offers the
possibility to install a subset of modules in the local machine, and to leverage
a Openstack environment able to perform the same operations of the
production-grade installation. Thus, we started configuring the Devstack
installation, installing first of all the Nova module, that is required to
perform computing operations such as the launch of virtual machines, and lately
we installed Tacker too.

@@

First of all, Devstack needs a machine with at least 16GB of ram and 50GB of HD
space. We decided to install these components:
\begin{itemize}
 \item Keystone: identity management module
 \item Object storage
 \item Compute
 \item Tacker: a VNF and NFVO (as explained), tested to see if Openstack was a
   suitable VIM too
\end{itemize}

@@

A short description about every installation/deploy method is briefly
illustrated.

@@

\vfill

@@

After the deployment of Openstack and Openbaton, the next logical step is to try
and make the component talk to each other. First, we tried a simple \verb!iperf!
data transmission with two VNFs deployed by Openbaton as a common
hypervisor-based Openstack instances, as explained in~\cite{openbatonIperf}.
With this test completed, we started experimenting Openstack so as to use Docker
containers instead of traditional hypervisor instances. When we started
developing this solution, we adopted the Zun module~\cite{zunmoduleos}, that
aims to smoothly integrate Docker in Openstack. Unfortunately, we were not able
to make it work properly. In particular, it was not possible to open
communications ports on the Docker containers, hence making it impossible for
the VNFs to communicate. On top of that, we found difficulties regarding the
integration of Docker images in TOSCA definitions and Openstack. These two
issues are related: the TOSCA definition does not have any keyword to
specificate in any way that the image used for the VNF definition is a Docker
image, and Openstack, with the Zun plugin, at the time of the implementation did
not performed the image pull from any Docker registry. The impossibility to
easily set the containers ports and the lack of some support between the TOSCA
definition and the Openstack Zun plugin pushed us to drop Openstack as a
possible solution and to start looking for another valid VIM component.

@@

After the failure to find a viable setting for Openstack, we decided to look
around for other solutions. We started studying proper Docker container
orchestrators that could offer all the features we needed: container lifecycle
management, scaling, healing and monitoring. We decided to pick two of the most
mature and famous ones: Docker Swarm and Kubernetes. The decision to study only
these two, relies on the big communities supporting them and the possibility to
easily find Open Source projects built on them that used the aforementioned
frameworks.

@@

First we started assesing the Docker Swarm orchestrator. Docker Swarm is a
product directly made by the Docker Inc., so it seamlessly integrates with other
Docker tools such as Docker Compose. The installation and set-up is pretty
straightforward, and the configuration is almost automatically handled by Docker
itself. The minimum number of nodes needed to create a proper Docker Swarm
cluster is three, where one must be a manager node (designed to orchestrate the
other nodes in the cluster), while the others can be simple workers. Multiple
managers can coexist in the same cluster, as long as only one effectively
operates. The others act like backup managers in case the one orchestrating goes
down for some reason. The networking is given out-of-the box, and the same holds
for distributed persistence. While this for most of the uses cases can be
considered a pro, in our specific situation the impossibility to tweak the
network configuration was a cons, since it was important for us to have the full
control over the network, especially regarding the SFC implementation, explained
better in Chapter~\ref{chap:vnf_ns_impl}.

\subsection{Kubernetes}
Kubernetes, as already described in the technology section, is an orchestrator
created by Google able to create a software abstraction from the underlying
virtualization technology. This enables Kubernetes to create an integration with
Docker. Kubernetes has an architecture based on plug-ins, giving the possibility
to the users to change part of Kubernetes itself, such as the networking routing
strategy or the persistence layer. With this flexibility Kubernetes offered to
us what Docker Swarm was not able to give. Additionally, we developed a custom
Domain Name System (DNS) lookup process and distribute other internal components
in order to avoid requests bottleneck (as explained better in
Section~\ref{chap:archimple:sec:secondattempt:sub:scalability}).

@@

Kubernetes does not require a minimum number of nodes in order to properly work.
Nonetheless, we decided to use a four node cluster configured in the following
way: three slaves nodes (also known as \emph{minions}) and one master node. The
Kubernetes version installed was the 1.11, enabling us to exploit new features
like a reworked DNS lookup system and new storage features\footnote{A full
  kubernetes log release can be found at:
  \url{https://kubernetes.io/blog/2018/06/27/kubernetes-1.11-release-announcement/}}.
For the initial deployment we relied on an internal Openstack instance installed
at the University of Padova made of four nodes with 4GB of RAM, 10GB of SSD
storage and 2vCPU on Ubuntu 16.04 machines (that we later swapped with CentOS
7). We performed the installation using the official Kubernetes repositories,
taking care of performing a accurated cluster configuration, as better explained
in Appendix~\ref{kubernetes}. After the initial installation, we obtained a
configuration similar to the one illustrated in
Figure~\ref{chap:archimpl:sec:secondattempt:img:k8scp}. In this figure is
possible to denote different components on the master node and on the minions,
which will be described.

@@

After the Kubernetes configuration and installation, we started working for a
viable integration between Kubernetes and Openbaton. Our first idea is the one
illustrated in Figure~\ref{chap:archimpl:sec:secondattempt:img:schema1}:
Openbaton, that usually handles the TOSCA definitions storing them in its
database, had to somehow convert these definition in a suitable YAML for
Kubernetes, that had in turn to elaborate it and schedule to one minion in the
cluster. The monitoring and management would remain untouched, as we planned to
create an image with the integrated EMS tools that provided a seamless
integration with Docker container. Under these settings, a key issue is the
conversion of TOSCA definitions in YAML ones. This topic was quite important
because the TOSCA definition does not really consider the possibility to have an
alternative virtualization system apart from the hypervisor one. In particular,
we found the Openbaton TOSCA definition to be virtual machine centric, without
providing the possibility of formulating a agnostic approach to virtualization
enabling container-based deployments. Nonetheless, we performed a mapping
between most of the TOSCA keywords with some of the YAML definitions, in a way
that a TOSCA configuration file was able to generate a valid YAML deployment.
After that, we started thinking who, in the current architecture, had the task
to perform this operation. Initially we thought that Kubernetes itself could
have been a perfect candidate. From a received TOSCA, a Kubernetes component
should have validated and translated the TOSCA definition to generate a YAML
suitable to be finally submitted to the Kubernetes API Server. Another feasible
approach could have been to create a Openbaton plugin in order to convert the
TOSCA definition before submitting it to Kubernetes for the deployment. This
solution had great advantages in terms of responsibility separation: in this way
only the MANO had to know, for a VIM based on Kubernetes, that a conversion was
needed. From a Kubernetes point-of-view, nothing would have changed, since all
the configuration conversion would have been performed by Openbaton.

@@

The impossibility to develop a custom VIM driver and VNFM manager, given the
resources of hand, led us to search for an alternative solution. Then, we
started to work directly on the VIM. The base approach was to create an API
Server which in future an Openbaton plugin could contact and simply deploy TOSCA
definitions, interacting with a set of RESTful APIs. We called this simple
backend \emph{Harbor}.

@@

Terminated the Harbor development, we ended up in a situation similar the one
depicted in Figure~\ref{chap:archimple:sec:secondattempt:img:attempt2v1harbor}.
It is possible to see how the Openbaton framework is not necessary anymore:
Harbor, in fact, ended up substituting it. We then started to think about the
challenges of integrating the SFC structure inside a Kubernetes cluster. We
identified the following challenges:
\begin{itemize}
\item whole system scalability: we want for every component to have the
  possibility of scaling during intense workloads, without the necessity of
  human intervention;
\item ingress and egress points: specularized entities of our framework handling
  SFC encapsulation/decapsulation session data;
\item dynamic packet routing: as formulated by one of our requirements, packet
  routing had to be dynamic. We immediately identified a possibility to create
  this functionality with the already-built features of Kubernetes;
\item a framework to automatically forward the data elaborated by the VNF: we
  identified the possibility to build a framework that automatically handled the
  packet receiving from another VNF, it properly called the function that had to
  elaborate the packet payload and then it forwarded the result to the next VNF;
\item connection status: having ingress and egress points implicates the
  necessity to save some persistent data in a distributed manner.
\end{itemize}
In the following, we provide some details on how we achieved the aforementioned
properties.

\subsubsection{Scalability considerations}
\label{chap:archimple:sec:secondattempt:sub:scalability}
When adopting Kubernetes, we realized the possibility to make every component
easily scalable. Indeed, all the following components can be described in
Kubernetes as a \verb!Service!. A \verb!Service! can have zero or many
\verb!Deployment!s, giving the possibility for it to scale under heavy
workloads. Contemplating for scalability meant that it was not possible to refer
to the components based on their IP, but we had to refer to their service name.
As it will be explained later, this introduced challenges especially for the
ingress and egress points of our network, that have to keep a connection with
the sender and the receiver. On top of that, the scalability and the required
DNS lookups meant a possible bottleneck. With a great number of components and
VNFs constantly querying the centralized Kubernetes DNS lookup system a
performance degradation could have easily happened. We decided to distribute the
DNS service too, integrating the CoreDNS Kubernetes default service with
\emph{Dnsmasq}, a DNS server able to perform query caching for faster name
resolution. We created a Dnsmasq \verb!DaemonSets!, which in turn deployed one
Pod with Dnsmasq running inside for every minion in the cluster. At the time of
the VNF launch, we set that the DNS resolution should pass first for the local
Dnsmasq DNS, and then for the Kubernetes CoreDNS service. In this way, we were
able to distribute the DNS query workload from the master, using a local caching
system to speed up the name resolution between services.

\subsubsection{Ingress and Egress: Ironhide}
As just said, the whole framework needs to have and entry and an exit point.
These entities, defined as \emph{Ingress} and \emph{Egress} points, are
provisioned by \emph{Ironhide}. Ironhide hides the whole packet processing to
the sender $S$ and the receiver $R$. When a client $S$ performs a request to
$R$, the packet actually goes through Ironhide, who has an internal classifier
that it is able to recognize the type of connection and to choose the right NS
the packet has to go through. This process is fundamental since it keeps the
connection between $S$ and $R$ up, for both TCP and UDP based communications.
Since the connection needs to be end-to-end, the state of the connections are
saved in the \emph{Roulette Controller} component.

@@

\subsubsection{Keeping the connection status}
Desining for scalability gave rise to new issues. The most important is keeping
up stateful connections, since the ingress and egress points could be more than
one container. With this consideration in mind, we started thinking at how and
where keeping information about the ingress container handling the connection
with the sender $S$ and about the egress container talking with the receiver
$R$. Since the Roulette component has a complete end-to-end topology view of the
system, we decided to store these additional information (we called it
\emph{endpoints}) in it.

@@

\subsection{Discussion}
At the end, we were able to create a scalable architecture that followed the
ETSI MANO standard, with Kubernetes as VIM and Harbor as MANO. We faced
additional challenges regarding the system scalability, that we solved using the
Roulette controller. We eased the creation of VNFs with Astaire, and we created
proper ingress and egress points with Ironhide, that supported scalability
issues and performed a protocol-based packet classification. Is possible to see
a final architectural schema in
Figure~\ref{chap:archimpl:sec:secondattempt:img:attempt2v1total}.
